\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-11-08 casestudies.tex
Update Information on AI example.

Our second case study is a mechanization of abstract interpreters
for simple imperative languages.
In addition to the metatheories, this case study produces abstract
interpreters that are directly runnable in Coq.

The code is organized into four families.
A base family, \lsti{Imp}, defines the abstract syntax of a while-language
with pure expressions and impure statements, via \lsti{FInductive}.
It also defines a big-step interpreter parameterized by a fuel value,
via \lsti{FRecursion}.
It contains \textasciitilde 200~LOC.

A second family, \lsti{ImpGAI}, extends \lsti{Imp} and contains
\textasciitilde 450~LOC.
It exports a generic framework for deriving provably sound
abstract interpreters.
The soundness theorem of this generic abstract interpreter, \lsti{analyze},
is with respect to the big-step interpreter, \lsti{eval}, inherited from \lsti{Imp}.
It states that the abstraction relation \lsti{RState} over a
concrete state~\lsti{S} and an abstract state~\lsti{absS} is preserved
by the analysis:

\begin{centered}
\begin{minipage}{.86\textwidth}
\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
\forall stmt fuel S absS, RState S absS -> RState (eval fuel stmt S) (analyze stmt absS)
\end{lstlisting}
\end{minipage}
\end{centered}

@@

\noindent
\lsti{analyze} is defined via \lsti{FRecursion}, and the soundness theorem
is proved via \lsti{FInduction}.
This family leaves fields representing the abstract domain, the
abstraction relation, monotonicity of transfer functions, etc.\ 
unspecified or unproven---a derived family can instantiate these
``parameters'' by overriding the fields (and also possibly extend the abstract syntax),
to create a sound, runnable abstract interpreter for a (possibly extended) while-language.

@@

Our implementation of family polymorphism retains the computation
ability of the host proof assistant;
the two abstract interpreters in \lsti{ImpTI} and \lsti{ImpCP} are
directly runnable in Coq.
We also extract them to OCaml and put them to test over queries.

>>>>

Our second case study is a mechanization of abstract interpreters
for simple imperative languages.
In addition to the metatheories, this case study produces abstract
interpreters that are directly ready for program extraction in Coq.

The code is organized into four families.
A base family, \lsti{Imp}, defines the abstract syntax of a while-language
with pure expressions and impure statements, via \lsti{FInductive}.
The semantic is formulated as a functional interpreter like a CEK machine\cite{felleisen1986control}, parameterized by a fuel value,
via \lsti{FRecursion}.
It contains \textasciitilde 200~LOC.

A second family, \lsti{ImpGAI}, extends \lsti{Imp} and contains
\textasciitilde 550~LOC.
It exports a generic framework for deriving
abstract interpreters with partial correctness.
The soundness theorem of this generic abstract interpreter, \lsti{analyze},
is with respect to the interpreter, \lsti{eval}, inherited from \lsti{Imp}.
It states that the abstraction relation \lsti{RState} over a
concrete state~\lsti{S} and an abstract state~\lsti{absS} is preserved
by the analysis:

\begin{centered}
\begin{minipage}{.86\textwidth}
\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
\forall stmt fuel S absS, RState S absS -> RState (eval fuel stmt S) (analyze fuel stmt absS)
\end{lstlisting}
\end{minipage}
\end{centered}

@@

\noindent
\lsti{analyze} is defined via \lsti{FRecursion}, and the soundness theorem
is proved via \lsti{FInduction}.
This family leaves fields representing the abstract domain, the
abstraction relation, monotonicity of transfer functions, etc.\ 
largely unspecified or unproven\EDJ{It is not very relaxed. Currently everything is using a dicionary as memory model and stick with that. The abstract value for each expression is the one *really* unspecified.}---a derived family can instantiate these
``parameters'' by overriding appropriate fields (and also possibly extend the abstract syntax),
to create a sound, runnable abstract interpreter for a (possibly extended) while-language.

@@

Our implementation of family polymorphism retains the computation
ability of the host proof assistant;
the two abstract interpreters in \lsti{ImpTI} and \lsti{ImpCP} are
directly ready for program extraction to OCaml and put them to test over queries.
